The role of advection versus diffusion in intraventricular transport

### 2025-02-03 paper/main.tex
Results: modify sections 2-5, add a reference

%% Some new commands
% Partial derivatives
\newcommand{\pdifft}[1]{\frac{\partial  #1}{\partial t}}
\newcommand{\pdiffx}[1]{\frac{\partial #1}{\partial x}}
\newcommand{\pdiffy}[1]{\frac{\partial #1}{\partial y}}
\newcommand{\pdiffxi}[1]{\frac{\partial #1}{\partial x_i}}
\newcommand{\pdiffxj}[1]{\frac{\partial #1}{\partial x_j}}
\newcommand{\secpdiffx}[1]{\frac{\partial^2 #1}{\partial x^2}}
\newcommand{\secpdiffxi}[1]{\frac{\partial^2 #1}{\partial x_i \partial x_i}}
\newcommand{\secpdiffxj}[1]{\frac{\partial^2 #1}{\partial x_j \partial x_j}}

@@

\subsection*{Quantities of interest}
We calculate the mean solute concentrations
\begin{equation}
    \overline{c}_i(t) = \frac{1}{|\Omega_i|}\int_{\Omega_i} c(\xx, t)\,\dx,
    \label{eq:c_mean_i}
\end{equation}
in the six regions of interest (ROIs) $\Omega_i$ ($i = 1, \dots, 6$) as functions of time, where $|\Omega_i|$ is the volume of the domain $\Omega_i$. The dynamics of ROI 1 follows directly from~\eqref{eq:injection_curve}. In addition, we report time-to-threshold: the time $\hat{t}_i$ when the mean concentration $\overline{c}_i$ first exceed a threshold value $\hat{c}_i$. We use $\hat{c}_i=0.25$ for ROI 1--4 and $\hat{c}_i=0.10$ for ROI 5--6.

@@

\subsection*{Ventricular solute transport is advection-dominated for larger molecules}
We next ask what is the impact of these CSF flow patterns and
magnitudes on solute transport generally within the brain ventricles?
To address this question, we simulate the distribution and evolution
of different solutes subject to photoconversion within a region of
interest in the dorsal region of the middle ventricle (ROI 1)
(\Cref{fig:fig2}\fixme{e/x}). In particular, we consider the spread of
a solute concentration with a given molecular diffusion coefficient,
corresponding to Exosomes ($D_1 = 1.63 \times 10^{-12}$ m$^2$/s),
STM+GFP ($D_2 = 5.75 \times 10^{-11}$ m$^2$/s), and Dendra2 ($D_3 =
1.15 \times 10^{-10}$ m$^2$/s). As expected, transport occurs more
rapidly with increasing diffusion coefficient. Interestingly, the
distribution patterns differ between the different-characteristics
solutes. For the smallest diffusion coefficient, representing
Exosomes, the advection induced by the CSF flow dominates diffusion
(\Cref{fig:fig2}a). The vortex structures of the velocity field
carry over into the concentration field, indicating that transport
along the streamlines is more rapid than the diffusion across the
streamlines. The balance between diffusion and advection shifts with
increasing diffusion coefficients, the CSF flow structures become more
obfuscated, and the solute spreads more uniformly throughout the
ventricular geometry (\Cref{fig:fig2}\fixme{x-y}). Interestingly, the
smaller solutes with higher diffusion coefficients spread more quickly
to regions distal to the photoconversion site (\Cref{fig:fig2}\fixme{z}).
\begin{figure}%[H]
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}
    \caption{Simulated distribution and evolution after photoconversion of Dendra ($D_1$), STM+GFP ($D_2$) and Exosomes ($D_3$).
    \mer{MER: I think you should show on the same color scale here in a) (the range of values lets you and then it is possible to compare). Split a into a, b, c. Show e) first (so that e) becomes a)}.
    \textbf{a.} Slices in the $xz$-plane ($y=0.140$ mm, the center of the geometry) showing the concentration field simulated with diffusion coefficients $D_1$, $D_2$ and $D_3$ for three time instants $t=45.0$ s, $t=112.6$ s, $t=450.5$ s.
    Note that the colorbars indicating the magnitude of $c$ have different limits for each column.
    \textbf{b--d.} Mean concentration $\overline{c}$ in each region of interest (ROI) as function of time for diffusion coefficient $D_1$, $D_2$ and $D_3$, respectively.
    \textbf{e.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{f.} The time $\hat{t}$ before the mean concentration $\overline{c}$ exceeds a threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}
    \label{fig:fig2}
\end{figure}

@@

Transport of solute is quantified by the profile of the mean concentrations $\overline{c}_i(t)$ in the regions of interest (ROIs) $\Omega_i$ as functions of time~(\Cref{fig:fig2}b--d).
An illustration of the ROIs is provided in \Cref{fig:fig2}e. The dynamics of ROI 1 follow from imposing~\Cref{eq:injection_curve}.
For ROIs 2--6, we observe more rapid concentration increases when the diffusion coefficient increases. 
For all three diffusion coefficients, the dynamics of ROI 6 is the slowest.
Region 5 exhibits a more rapid increase in mean concentration, and an even faster increase is observed for ROIs 2--4.
The profiles of the latter three regions are similar, displaying only small differences in all of the three plots.
\lyng{HH: Add some quantification in terms of percentages, e.g. final values of $\overline{c}_i$ for the different $D_i$?}
In region 6, the threshold is never exceeded for $D_1$, with the mean concentration close to zero at the end of the simulation. 
To quantify the balance of advection versus diffusion as transport mehcanisms on a global scale, we calculate the Péclet numbers $\mathrm{Pe}_1=519$, $\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.
Since $\mathrm{Pe}>1$, this indicates that advection is the dominant transport mechanism for the molecules considered.

\subsection*{Comparing experimental and computational results}
Transport of photoconverted proteins is quantified by the change in fluorescence intensity $\Delta F=(F(t)-F_0)/F_0$ over time~(\Cref{fig:figure3_compare_exp_and_sim_control}a).
The solid lines are the mean values of $\Delta F$ for the fish cohort, with shaded regions covering one standard deviation.
The first region of interest (ROI 1) is the laser-targeted region. Qualitatively, the fluorescence intensity change in this region resembles intensity dynamics observed in experiments that studied the photoconversion properties of Dendra2 fluorescent proteins~\cite{Makarov2014Steady-stateDendra2}. The transport to the other regions can be separated into two groups: the fluorescence intensity change $\Delta F$ for ROIs 2--4 in the middle ventricle are all similar, whereas the more distal regions ROI 5 and 6 (anterior and posterior ventricle, respectively) exhibit a slower increase in $\Delta F$. In ROIs 2--4, the $\Delta F$ curves are concave, similar to ROI 1, while in ROIs 5 and 6 the curves are convex. Although the fluorescence intensity increases throughout all of the experiments, the maximum $\Delta F$ is not necessarily attained at the final time of the experiments. At the final time, the mean fluorescence intensities in regions 2--4 all exceed 0.60 (0.63, 0.71, 0.67, respectively) of the final value $F_0$ in ROI 1, whereas ROIs 5 and 6 reach mean values of 0.29 and 0.18, respectively.

% Final Delta F values
%0.6322650156762044 ROI 2
%0.7059449124175479 ROI 3
%0.6694500367160006 ROI 4
%0.28986700094045437 ROI 5
%0.18155055624229066 ROI 6

The experiment of photoconverting Dendra2 was simulated by solving the advection-diffusion equation with the diffusion coefficient $D_3$. We quantify the simulated transport with the time evolution of mean concentrations $\overline{c}$~(\cref{fig:figure3_compare_exp_and_sim_control}, panel \textbf{b}). Imposing the concentration in ROI 1 mimics the fluorescence intensity change $\Delta F$ in ROI 1. Since the initial condition of the simulation is a mean concentration of zero, the mean concentration $\overline{c}$ can be interpreted as the change in mean concentration and is thus a similar metric to $\Delta F$.

Comparing the simulation results with experimental results, transport dynamics of regions 2--6 are faster in the simulation. We observe higher final values of $\overline{c}$ compared to the final values of $\Delta F$ in the experiments. In simulations, the three $\overline{c}$ profiles of ROIs 2--4 behave similar as the $\Delta F$ profiles in these regions, although the orders of ROI 3 and 4 are opposite in the simulations as compared to the experiments. For all regions, the maximum mean concentration is attained at the final time of the simulations. The final value of $\overline{c}$ in regions 2--4 were 0.76, 0.72 and 0.82, respectively, compared to 0.63, 0.71 and 0.67 for $\Delta F$. As in the experiments, the transport to regions 5 and 6 is slower than the transport to the middle ventricle sites, though the final values of $\overline{c}$ are higher than those of $\Delta F$. The anterior ventricle (region 5) has a final $\overline{c}$ value of 0.56 (final $\Delta F= 0.29$), while in the posterior ventricle (region 6) the final value is 0.35 (final $\Delta F= 0.18$). Qualitatively, the $\overline{c}$ profile in the posterior ventricle compares with $\Delta F$ in this region, displaying a convex curve that has an approximately linear increase after an initial nonlinear increase. The anterior ventricle (region 5), on other hand, has an S-shaped $\overline{c}$ curve in the simulations, which differs from the linear $\Delta F$ curve in the experiments. 
% Final mean c values ROIs 2-6: 0.7637806  0.72025859 0.81903216 0.56204164 0.34926411
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure3_compare_experiments_and_simulations.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $\Delta F$ over time in each region of interest (ROI) observed during experiments of photoconversion of Dendra2 proteins in ROI 1. \textbf{b.} Mean concentration $\overline{c}$ as function of time in each ROI in a simulation of Dendra2 protein transport with the diffusion coefficient $D_3$. \textbf{c.} Time to $\Delta F$ reaches a threshold value of 0.25 in one of the animals of the dataset presented in panel \textbf{a}. \textbf{d.} Time to $\overline{c}$ reaches a threshold value of 0.25 for a transport simulation with diffusion coefficient $D_3$.}
    \label{fig:figure3_compare_exp_and_sim_control}
\end{figure}

We qualitatively compared fluorescence intensities and mean concentrations with the time $\hat{t}$ to exceed a threshold value of 0.25~(\cref{fig:figure3_compare_exp_and_sim_control}, panels \textbf{c} and \textbf{d}). For $\Delta F$, we determined this time-to-threshold for the same downsampled dataset that was used to calculate $\Delta F$. In the simulations, we determined the time for $\overline{c}$ to reach the threshold value in each degree of freedom of the concentration finite element function. We have excluded the posterior-most regions of the geometries in~\cref{fig:figure3_compare_exp_and_sim_control}, because the values of $\Delta F$ and $c$ never exceeded the threshold value of 0.25 in these regions.

\subsection*{Absence of ciliary motion affects local protein distribution}
We compare the time evolution of fluorescence intensity in control and \emph{schmalhans}~(\emph{smh}) mutant zebrafish, plotting the means of $\Delta F$ across the cohorts, shaded with one standard deviation regions~(\cref{fig:figure4_compare_control_mutant}, panel \textbf{a}). Additionally, we determine the time needed to reach a threshold fluorescent intensity change in each region of interest (ROI) for each fish~(\cref{fig:figure4_compare_control_mutant}, panel \textbf{b}). It is important to note that this threshold was set differently for two groups of regions: the $\Delta F$ threshold value was set to 0.25 for ROIs 1--4, whereas for ROIs 5 and 6 the threshold value was 0.10. This choice was made because of the slower dynamics of ROIs 5 and 6. For each region of interest, the individual data points of the times to threshold are presented together with the distributions and the means of the data points. Based on the times to reach a threshold value of $\Delta F$, we test the null hypothesis that the control and mutant photoconversion data are from the same population (or equivalently, have equal distributions). We report the $p$-value for this hypothesis calculated using a two-sided Mann-Whitney U test that included a continuity correction, with a confidence level of 95\%.

In ROI 2, there is a statistically significant slower fluorescence intensity increase in the \emph{smh} mutant fish compared to controls~(\cref{fig:figure4_compare_control_mutant}, panel~\textbf{b}). Compared to the control dataset, the mean time-to-threshold in ROI 2 is 45\% higher for the mutant dataset. In the other ROIs, the differences in the experimental results for controls and mutants are not significantly different. We observe that, both within the control and within the mutant dataset, there are considerable individual differences in the times to reach the threshold.

To simulate protein transport in mutant fish, we used flow model II (only cardiac motion as flow mechanism, no cilia) to model the ciliary dysfunction of the \emph{smh} mutants. We visualize the time evolution of the mean concentration $\overline{c}$ in each ROI for both model II and the baseline model (model 0), in addition to a similar time-to-threshold $\hat{t}$ as computed for the experimental data~(\cref{fig:figure4_compare_control_mutant}, panel~\textbf{c}). As for the experiments, we use a threshold value of $0.25$ in regions 2--4 and a value of $0.10$ for regions 5 and 6 to calculate $\hat{t}$. For ROI 2, we observe the same dynamics in simulations as in experiments: there is a slower increase in mean concentration in ROI 2 compared to the baseline model. The difference is even greater for the simulations, with the time-to-threshold in ROI 2 for model II 90\% higher than for model 0. Additionally, ROIs 4 and 5 exhibit very different $\overline{c}$ profiles for the two different flow models. The difference is less significant for ROI 3, and for the posterior ventricle (region 6) the profiles are very similar. 
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $F$ over time in each region of interest (ROI) for control (black lines) and \emph{smh} mutant (red lines) fish. 
    \textbf{b.} Times when $\Delta F$ first exceeds a threshold value, equal to 0.25 (ROIs 2-4) or 0.10 (ROIs 5, 6), for control (gray dots) and \emph{smh} mutant (red dots) fish. The thick black lines denote mean values of the datasets, and $p$-values were calculated using a Mann-Whitney U test with a confidence level of 95\%.
    \textbf{c.} Mean concentration $\overline{c}$ as function of time in each ROI, simulated using the baseline model (black lines) and the cardiac-only model (red lines). The horizontal gray lines mark the threshold value $\hat{c}$, and the time $\hat{t}$ when the threshold is first exceeded is reported in the plot legends. Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have different limits in each plot.}
    \label{fig:figure4_compare_control_mutant}
\end{figure}

\subsection*{Loss of cilia can slow down transport}
Considering removal of cilia in three different regions through modified flow boundary conditions, we simulate CSF flow modeled by the Stokes equations and use the resulting velocity field in transport simulations of Dendra2. The maximum velocity observed in the simulations when removing telencephalic, ventral diencephalic and dorsal diencephalic cilia were, respectively, 27.7 $\mu$m/s, 27.2 $\mu$m/s, and 8.0 $\mu$m/s. We remind the reader that the maximum velocity observed for the baseline model was 27.7 $\mu$m/s.

Upon removal of telencephalic cilia, the evolution of the mean concentration $\overline{c}$ with time is essentially unaltered~(\cref{fig:figure5_compare_cilia_modifications}, panel \textbf{b}). Without ventral cilia in the diencephalic ventricle, $\overline{c}$ increases slightly slower in region of interest (ROI) 3 compared to the baseline model. Furthermore, there is a higher mean concentration in the anterior ventricle (ROI 5) and a lower mean concentration in the posterior ventricle (ROI 6) in the absence of ventral cilia when comparing to the baseline model, meaning more solute has spread towards the anterior ventricle and less solute has spread towards the posterior ventricle.

Changes in the transport dynamics are most prominent when removing the dorsal cilia in the diencephalic ventricle. For this setup, all ROIs (except ROI 1, for which we impose the concentration dynamics) exhibit significant changes in the dynamics of $\overline{c}$. Compared to the baseline model results, there is a slower increase in, and lower final values of, the mean concentration in middle and anterior ventricles (ROIs 2--5). In contrast, the transport to the posterior ventricle (ROI 6) is more rapid upon removal of the dorsal diencephalic ventricle cilia. It is evident from the final values of $\overline{c}$ that there is less total mass added to the system in the absence of dorsal cilia. \lyng{HH:This is due to the lower velocities in the vicinity of ROI 1, resulting in less ability to carry away mass from the volume where the concentration is imposed. \emph{Is this correct?}}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{Caption}
    \label{fig:figure5_compare_cilia_modifications}
\end{figure}

\subsection*{Modification of ventricles geometry impacts protein distribution}

The brain ventricular geometry varies significantly across zebrafish individuals, both healthy and sick animals. Motivated by an interest in how these variations can impact solute transport in the brain ventricles, we study the impact on simulation results when modifying the computational geometry and mesh in four different scenarios. Because our computational mesh is inflated by the dye-injection performed before image segmentation, we study four different mesh versions where we shrink: the fore-mid brain connection; the middle ventricle; the mid-hind brain connection; and all of the three above simultaneously. We used Blender~\cite{Community2018BlenderPackage} to modify the brain ventricles geometry.

We simulate flow and transport in four alternative brain ventricles meshes: one where the connection between the anterior and middle ventricles is constricted; one where the connection between the middle and posterior ventricles is constricted; one where the middle ventricle is shrunk in the lateral direction; the combination of the above three modifications~(\cref{fig:figure6_sim_results_compare_geometries}, panel \textbf{a}). We simulate CSF flow in the modified geometries with the baseline flow model and use the resulting velocity field to simulate transport of Dendra2.

@@

\lyng{HH: Quantify differences in terms of percentage changes in the final mean concentration values, maybe with a bar plot?}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{Caption. \mk{MK: How do the different geometries compare in terms of volume, areas where cilia are active?} \lyng{HH: Good question, will add this information.}}
    \label{fig:figure6_sim_results_compare_geometries}
\end{figure}

\cleardoublepage

@@

\subsection*{The roles of diffusion and advection in transport of photoconverted proteins}
We quantified transport in terms of fluorescence intensity change $\Delta F$ (experiments) and mean concentration $\overline{c}$ (simulations) as functions of time. For both measures, we observed that transport dynamics in the middle ventricle where Dendra2 fluorescent protein was photoconverted contrasted with the dynamics of the more distal regions of the anterior and posterior ventricles. In experiments, regions distal to the photoconversion site showed an approximately linear increase in $\Delta F$, probably because advection is absent in the transport to the distal regions, such that diffusion is the sole mode of transport. In regions proximal to photoconversion, advection and diffusion occur simultaneously due to the presence of cilia, leading to concave profiles for $\Delta F$. The convex diffusion-only behavior was observed in the mean concentration $\overline{c}_6$ profile of the posterior ventricle in the simulations, but $\overline{c}_5$ was concave for the anterior ventricle, opposed to the experimental results \lyng{HH: \emph{Is this the correct interpretation of convex vs. concave curves?} Include some referencing to textbook solutions of the diffusion and advection equations to explain this convex vs. concave? E.g. semi-infinite slab or boundary layer solutions to inflow/outflow problem.}. Based on the geometry modification study, the reason for this difference in behavior might be that the duct between the anterior and middle ventricle is inflated in our computational mesh, as a result of the dye-injection performed before imaging~\lyng{Cite this?}. This inflation allows for effective cilia-mediated advection from the middle ventricle to the anterior ventricle.

>>>>

%% Some new commands
% Math
\newcommand{\cbar}{\overline{c}}

% Partial derivatives
\newcommand{\pdifft}[1]{\frac{\partial  #1}{\partial t}}
\newcommand{\pdiffx}[1]{\frac{\partial #1}{\partial x}}
\newcommand{\pdiffy}[1]{\frac{\partial #1}{\partial y}}
\newcommand{\pdiffxi}[1]{\frac{\partial #1}{\partial x_i}}
\newcommand{\pdiffxj}[1]{\frac{\partial #1}{\partial x_j}}
\newcommand{\secpdiffx}[1]{\frac{\partial^2 #1}{\partial x^2}}
\newcommand{\secpdiffxi}[1]{\frac{\partial^2 #1}{\partial x_i \partial x_i}}
\newcommand{\secpdiffxj}[1]{\frac{\partial^2 #1}{\partial x_j \partial x_j}}

@@

\subsection*{Quantities of interest}
We calculate the mean solute concentrations
\begin{equation}
    \cbar_i(t) = \frac{1}{|\Omega_i|}\int_{\Omega_i} c(\xx, t)\,\dx,
    \label{eq:c_mean_i}
\end{equation}
in the six regions of interest (ROIs) $\Omega_i$ ($i = 1, \dots, 6$) as functions of time, where $|\Omega_i|$ is the volume of the domain $\Omega_i$. The dynamics of ROI 1 follows directly from~\eqref{eq:injection_curve}. In addition, we report time-to-threshold: the time $\hat{t}_i$ when the mean concentration $\cbar_i$ first exceed a threshold value $\hat{c}_i$. We use $\hat{c}_i=0.25$ for ROI 1--4 and $\hat{c}_i=0.10$ for ROI 5--6.

@@

\subsection*{Ventricular solute transport is advection-dominated for larger molecules}
We next ask what is the impact of these CSF flow patterns and
magnitudes on solute transport generally within the brain ventricles?
To address this question, we simulate the distribution and evolution
of different solutes subject to photoconversion within a region of
interest in the dorsal region of the middle ventricle (ROI 1)
(\Cref{fig:fig2}\fixme{e/x}). In particular, we consider the spread of
a solute concentration with a given molecular diffusion coefficient,
corresponding to Exosomes ($D_1 = 1.63 \times 10^{-12}$ m$^2$/s),
STM+GFP ($D_2 = 5.75 \times 10^{-11}$ m$^2$/s), and Dendra2 ($D_3 =
1.15 \times 10^{-10}$ m$^2$/s). As expected, transport occurs more
rapidly with increasing diffusion coefficient. Interestingly, the
distribution patterns differ between the different-characteristics
solutes. For the smallest diffusion coefficient, representing
Exosomes, the advection induced by the CSF flow dominates diffusion
(\Cref{fig:fig2}a). The vortex structures of the velocity field
carry over into the concentration field, indicating that transport
along the streamlines is more rapid than the diffusion across the
streamlines. The balance between diffusion and advection shifts with
increasing diffusion coefficients, the CSF flow structures become more
obfuscated, and the solute spreads more uniformly throughout the
ventricular geometry (\Cref{fig:fig2}\fixme{x-y}). Interestingly, the
smaller solutes with higher diffusion coefficients spread more quickly
to regions distal to the photoconversion site (\Cref{fig:fig2}\fixme{z}).
\begin{figure}%[H]
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}
    \caption{Simulated distribution and evolution after photoconversion of Dendra ($D_1$), STM+GFP ($D_2$) and Exosomes ($D_3$).
    \mer{MER: I think you should show on the same color scale here in a) (the range of values lets you and then it is possible to compare). Split a into a, b, c. Show e) first (so that e) becomes a)}.
    \textbf{a.} Slices in the $xz$-plane ($y=0.140$ mm, the center of the geometry) showing the concentration field simulated with diffusion coefficients $D_1$, $D_2$ and $D_3$ for three time instants $t=45.0$ s, $t=112.6$ s, $t=450.5$ s.
    Note that the colorbars indicating the magnitude of $c$ have different limits for each column.
    \textbf{b--d.} Mean concentration $\cbar$ in each region of interest (ROI) as function of time for diffusion coefficient $D_1$, $D_2$ and $D_3$, respectively.
    \textbf{e.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{f.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}
    \label{fig:fig2}
\end{figure}

@@

Transport of solute is quantified by the profile of the mean concentrations $\cbar_i(t)$ in the regions of interest (ROIs) $\Omega_i$ as functions of time~(\Cref{fig:fig2}b--d).
An illustration of the ROIs is provided in \Cref{fig:fig2}e. The dynamics of ROI 1 follow from imposing~\Cref{eq:injection_curve}.
For ROIs 2--6, we observe more rapid concentration increases when the diffusion coefficient increases. 
For all three diffusion coefficients, the dynamics of ROI 6 is the slowest.
Region 5 exhibits a more rapid increase in mean concentration, and an even faster increase is observed for ROIs 2--4.
The profiles of the latter three regions are similar, displaying only small differences in all of the three plots.
\lyng{HH: Add some quantification in terms of percentages, e.g. final values of $\cbar_i$ for the different $D_i$?}
In region 6, the threshold is never exceeded for $D_1$, with the mean concentration close to zero at the end of the simulation. 
To quantify the balance of advection versus diffusion as transport mehcanisms on a global scale, we calculate the Péclet numbers $\mathrm{Pe}_1=519$, $\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.
Since $\mathrm{Pe}>1$, this indicates that advection is the dominant transport mechanism for the molecules considered.

\subsection*{Validation of the simulations with experimental data}
To validate our computational model, we performed photoconversion experiments in transgenic zebrafish larvae expressing a secreted Dendra2 in their brain ventricles.
Our photoconversion protocol consisted of the acquisition of a baseline, before localized exposure to a UV laser (at the ROI 1 location)
that converts Dendra2 from a green to a red emitting fluorescent protein. 
To quantify the transport of photoconverted proteins, we measured and calculated the change in fluorescence intensity $\Delta F = (F(t)-F_0)/F_0$
over time (\Cref{fig:figure3_compare_exp_and_sim_control}a, $F_0$ being the average fluorescence at baseline).

We obtained very similar kinetics curves (\Cref{fig:figure3_compare_exp_and_sim_control}a, b)
and time-to-threshold values (\Cref{fig:figure3_compare_exp_and_sim_control}c, d) when comparing our
experiments (\Cref{fig:figure3_compare_exp_and_sim_control}a, c) with
simulation (\Cref{fig:figure3_compare_exp_and_sim_control}b, d).
Importantly, since the initial condition of the simulation is a mean concentration of zero,
the mean concentration $\cbar$ can be interpreted as the change in mean concentration and is thus a similar metric to $\Delta F$.
The final values of $\cbar$ were 0.76, 0.72, 0.82, 0.56 and 0.35 respectively,
compared to 0.63, 0.71, 0.67, 0.29 and 0.18 for $\Delta F$.
Notwithstanding similarities, we observed small differences between the experiments and simulation.
For instance, transport dynamics in ROIs 2--6 were faster in the simulation and reached higher final values.
Moreover, the $\cbar$ profiles of ROIs 3--4 were opposite order in the simulations as compared to the experiments.
Despite observing small differences, the \emph{in vivo} experiments and
simulations were similar enough to convincingly validate the model.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure3_compare_experiments_and_simulations.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $\Delta F$ over time in each region of interest (ROI)
    observed during experiments of photoconversion of Dendra2 proteins in ROI 1.
    \textbf{b.} Mean concentration $\cbar$ as function of time in each ROI in a simulation of Dendra2 protein transport with
    the diffusion coefficient $D_3$.
    \textbf{c.} Time to $\Delta F$ reaches a threshold value of 0.25 in one representative zebrafish larva.
    Time to $\cbar$ reaches a threshold value of 0.25 for a transport simulation with diffusion coefficient $D_3$.
    We have excluded the posterior-most regions of the geometries,
    because the values of $\Delta F$ and $\cbar$ never exceeded the threshold value of 0.25 in these regions.}
    \label{fig:figure3_compare_exp_and_sim_control}
\end{figure}
% Final Delta F values
%0.6322650156762044 ROI 2
%0.7059449124175479 ROI 3
%0.6694500367160006 ROI 4
%0.28986700094045437 ROI 5
%0.18155055624229066 ROI 6
% Final mean c values ROIs 2-6: 0.7637806  0.72025859 0.81903216 0.56204164 0.34926411

\subsection*{Absence of ciliary motion affects local protein distribution}
To assess the impact of cilia motion on solute distribution, we measured the movement of photoconverted Dendra2 signal
in control and motile-cilia mutant zebrafish (\emph{schmalhans} (\emph{smh})).
By plotting the means of $\Delta F$ (\Cref{fig:figure4_compare_control_mutant}a) and the time to threshold
across the control and mutant cohorts (\Cref{fig:figure4_compare_control_mutant}b), we identified a significant
difference in Dendra2 distribution in ROIs 2 and 4, but not in ROIs 3, 5, and 6. In ROIs 2 and 4,
which measure Dendra2 distribution downstream of the photoconversion site,
we observed a slower increase of fluorescence intensity,
resulting in a $\sim 45\%$ higher mean time-to-threshold in the \emph{smh}
mutant compared to controls (\Cref{fig:figure4_compare_control_mutant}b).

Next, to identify whether our model would recapitulate these findings,
we simulated protein transport by using the only-cardiac/no-cilia flow model 
and compared it with the baseline model (both cilia and cardiac motion) (\Cref{fig:figure4_compare_control_mutant}c).
Similar to our experiments, we observed slower dynamics in simulations for ROIs 2 and 4.
Interestingly, the difference was even greater for the simulations,
with the time-to-threshold in ROI 2 being 90\% higher in absence of cilia-mediated flow.
Moreover, ROI 5 exhibits very different $\cbar$ profiles for the two configurations in our computational model,
but not in our experimental data.
Altogether, our findings identify that diffusion plays an important role for solute movement to regions upstream in the cilia flow fields or to distant ventricles.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $F$ over time in each region of interest (ROI) for control (black lines) and \emph{smh} mutant (red lines) fish. 
    \textbf{b.} Times when $\Delta F$ first exceeds a threshold value, equal to 0.25 (ROIs 2-4) or 0.10 (ROIs 5, 6), for control (gray dots) and \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets, and $p$-values were calculated using a Mann-Whitney U test with a confidence level of 95\%.
    \textbf{c.} Mean concentration $\cbar$ as function of time in each ROI, simulated using the baseline model (black lines) and the cardiac-only/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have different limits in each plot.}
    \label{fig:figure4_compare_control_mutant}
\end{figure}

\subsection*{Loss of cilia motility in specific subpopulation slows down transport}
We previously identified three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the role of these lineages on CSF transport, we modified flow boundary conditions
in the three different ciliated regions (\Cref{fig:figure5_compare_cilia_modifications}a) and simulated
CSF flow using the Stokes equations. We identified that the velocities were most strongly
affected when removing cilia in the dorsal diencephalic cilia,
as the maximum velocity dropped from 27.7 $\mu$m/s to 8.0 $\mu$m/s.
In contrast, removing telencephalic and ventral diencephalic had minimalno to little impact and resulted in maximum velocities of 27.7 $\mu$m/s or 27.2 $\mu$m/s, respectively.

With the resulting velocity fields, we simulated transport of Dendra2.
Changes in the transport dynamics were most prominent when removing the dorsal cilia
in the diencephalic ventricle, which produces the strongest flow at this developmental stage.
All ROIs (except ROI 1, where we locally released solutes) exhibit significant changes
in the dynamics of $\cbar$ (\Cref{fig:figure5_compare_cilia_modifications}b).
Compared to the baseline model results, we observed a slower increase and lower final concentration
in the middle and anterior ventricles (ROIs 2--5),
and a faster transport to the posterior ventricle (ROI 6).
In contrast, removal of telencephalic cilia did not alter the evolution of the
mean concentration $\cbar$ with time (\Cref{fig:figure5_compare_cilia_modifications}b).
Finally, without ventral cilia in the diencephalic ventricle,
more solute has spread towards the anterior ventricle (ROI 5) and
less solute has spread towards the posterior ventricle (ROI 6).

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{Caption}
    \label{fig:figure5_compare_cilia_modifications}
\end{figure}

\subsection*{Modification of ventricles geometry impacts protein distribution}
We simulate flow and transport in four alternative brain ventricles meshes:
one where the connection between the anterior and middle ventricles is constricted;
one where the connection between the middle and posterior ventricles is constricted;
one where the middle ventricle is shrunk in the lateral direction;
the combination of the above three modifications~(\Cref{fig:figure6_sim_results_compare_geometries}a).
We simulate CSF flow in the modified geometries with the baseline flow model and use
the resulting velocity field to simulate transport of Dendra2.

@@

\lyng{HH: Quantify differences in terms of percentage changes in the final mean concentration values, maybe with a bar plot?}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{Caption. \mk{MK: How do the different geoemtries compare in terms of volume, areas where cilia are active?} \lyng{HH: Good question, will add this information.}}
    \label{fig:figure6_sim_results_compare_geometries}
\end{figure}

@@

\subsection*{The roles of diffusion and advection in transport of photoconverted proteins}
We quantified transport in terms of fluorescence intensity change $\Delta F$ (experiments) and mean concentration $\cbar$ (simulations) as functions of time. For both measures, we observed that transport dynamics in the middle ventricle where Dendra2 fluorescent protein was photoconverted contrasted with the dynamics of the more distal regions of the anterior and posterior ventricles. In experiments, regions distal to the photoconversion site showed an approximately linear increase in $\Delta F$, probably because advection is absent in the transport to the distal regions, such that diffusion is the sole mode of transport. In regions proximal to photoconversion, advection and diffusion occur simultaneously due to the presence of cilia, leading to concave profiles for $\Delta F$. The convex diffusion-only behavior was observed in the mean concentration $\cbar_6$ profile of the posterior ventricle in the simulations, but $\cbar_5$ was concave for the anterior ventricle, opposed to the experimental results \lyng{HH: \emph{Is this the correct interpretation of convex vs. concave curves?} Include some referencing to textbook solutions of the diffusion and advection equations to explain this convex vs. concave? E.g. semi-infinite slab or boundary layer solutions to inflow/outflow problem.}. Based on the geometry modification study, the reason for this difference in behavior might be that the duct between the anterior and middle ventricle is inflated in our computational mesh, as a result of the dye-injection performed before imaging~\lyng{Cite this?}. This inflation allows for effective cilia-mediated advection from the middle ventricle to the anterior ventricle.
